\documentclass{article}
\input{settings}
\newdate{date}{25}{09}{2023}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin now!
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title Page
\input{entry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Start the assignment now
% Some extra information!
\newpage

\addcontentsline{toc}{section}{Background}
\section*{BACKGROUND}
You will analyze a subset of the SPEC 2006 benchmark suite compiled for the \texttt{ia32} (32-bit Intel architecture) ISA. Specifically, you will instrument each SPEC 2006 application binary with PIN and extract the percentages of various different instruction types. You will also use a very simple cycle accounting model to calculate the CPI for each application. Additionally, you will calculate the instruction and data footprints of each application. Finally, you will understand a few properties of the \texttt{ia32} ISA.

In this assignment, we will be interested only in the following instruction types. At a very high level, we have two instruction types.

\vspace{10pt}
\noindent
Type A: Instructions with no memory operand \\
Type B: Instructions with at least one memory operand
\vspace{10pt}

You should only instrument the instructions that actually execute i.e., have true predicates. As a result, you should always use $\texttt{INS\_InsertPredicatedCall}$ method when classifying instructions.

For Type B Instructions, Each memory operand in such an instruction can be read and/or written to.

We will view each such load or store operation within a type B instruction as a separate micro-instruction and count it as a separate instruction.

For example, a type B instruction may have two load operations and a store operation. Further, each such load or store operation may access an arbitrary amount of data. However, in a 32-bit processor, it is not possible to transfer more than 32 bits of data in one shot. As a result, we will further break down each memory operation into several memory accesses of size at most 32 bits.

Finally, the instruction must be using these memory operands to carry out some operation of type A. This operation should be counted as a separate type A instruction and should also be categorized into one of the fifteen categories mentioned above.

Let us take an example of an x86 instruction that has three memory operands, two of which are loads and one is a store. The loads access one and ten bytes, respectively. The store accesses eleven bytes. This instruction will be accounted as follows.

\vspace{10pt}
\noindent
\texttt{Number of loads:   4}\\
\texttt{Number of stores:   3}\\
\texttt{Number of type A instruction:  1}
\vspace{10pt}

\noindent
Total instruction count increases by eight, whenever such an instruction is encountered.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Tasks pages begin
\newpage
\addcontentsline{toc}{section}{Tasks}
\section*{TASKS}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% New problem
\input{problem1}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% New problem
\input{problem2}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% New problem
\input{problem3}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% New problem
\input{problem4}
\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Complete the assignment now
\noindent

\section*{Sidenotes}
\begin{itemize}
    \item \texttt{TARGE\_IA32} is a preprocessor directive that is \texttt{true} when building for \texttt{ia32}
    \item All instructions that invoke the instrumentation routine are always classified at least as one of the Type-A. They can then be categorised in Type-B as well if they contain memory operands, and each 32B memory dealt with is treated as yet another purely Type-B instruction.
    \item There happens no categorisation of Type-A/B till the fast-forwarded amount of actual instructions are executed.
\end{itemize}

\begin{table}[H]
    \centering
    \begin{tabularx}{\linewidth}{|X|X|}
        \hline
        32b & 32 bits \\
        \hline
        32B & 32 Bytes \\
        \hline
        (B) & in Billions \\
        \hline
        \#Instructions & Number of Instructions \\
        \hline
    \end{tabularx}
    \caption{Abbreviations}
    \label{tab:abbreviations}
\end{table}

Please find more information regarding the submission in \texttt{README.md} in the submission directory.
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
